\Large \bf Cyclone: Warp Speed Replication for Key Value Stores

### 2017-07-21 paper/paper.tex
Updates to paper.
Switch to ASPLOS format.

\documentclass[10pt, preprint, nonatbib]{sigplanconf}
\usepackage{times,epsfig,endnotes,graphicx, verbatim}
\usepackage{datetime, url, pdfpages, parskip}

\conferenceinfo{SOSP'17}{October 29--31, 2017, Shanghai, China}
\copyrightyear{2017}

% These only appear when the 'preprint' option is specified.
% Enabling these will cause the first page of the document to fail the 
% format check on HotCRP :-(
%\titlebanner{Under submission to SOSP 2017 - do not cite or distribute}
%\preprintfooter{Draft of {\currenttime}, \today{}}

% No date in title area.
\date{}

% Paper number and no. of pages as author
\authorinfo{Paper \textbf{\#224}}{13 pages}

@@

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf Cyclone: Warp Speed Replication for Key Value Stores}

\maketitle

@@

\begin{abstract}
Persistent key value stores are rapidly becoming an important component of many
distributed data serving solutions with innovations targeted at taking advantage
of growing flash speeds. Unfortunately their performance is hampered by the need
to maintain and replicate a write ahead log to guarantee availability in the
face of machine and storage failures. Cyclone offers a solution to this problem
by leveraging a small amount of directly attached non-volatile memory (NVM) such
as NVDIMMs to maintain a two level log: the upper level in a small amount of NVM
draining into a larger lower level log maintained on flash. Replicating the
smaller log kept in NVMs transforms the replication problem into a packet
switching problem as the leader sends the exact same packet to all follower
replicas. We solve this problem by using the RAFT protocol implemented
as a software multicast module on the Data Plane Development Kit (DPDK). In addition,
we leverage the observation that key value store interfaces are commutative for
operations to different keys to scale our two level log horizontally into a
number of physical two level logs. A hash of the key is used to choose
the physical log. Each log is replicated by an independent instance of RAFT,
thereby leveraging the scalability afforded by classic software packet switching
when using multiple CPU cores. Cyclone is able to replicate millions of small
updates a second using only commodity 10 gigabit ethernet adapters, improving
the performance and availability of RocksDB, a popular persistent key value
store by an order of magnitude when compared to its own write ahead log without
replication.
\end{abstract}

@@

These key value stores however tend to ignore an important component: the write
ahead log. A machine or storage failure leading to total loss or temporary
unavailability of data is unacceptable in many internet scale services where
high availabilty and revenue are often interconnected. Key value stores
therefore usually incorporate support for a write ahead log that if replicated
and kept durable for every appended update provides the necessary high
availability. Unfortunately the write ahead is a performance achilles heel for
these systems - eclipsing much of the work on improving the performance of the LSM
component. To illustrate the impact of the write ahead log, consider
Figure~\ref{fig:problem}. The line marked 'Rocksdb' shows the performance of
Rocksdb without the write ahead log. The performance when persisting the write
ahead log without replicating it is shown as the line marked 'rocksdb/WAL'. The
line marked 'rocksdb/3 way rep.' is for simply replicating the log three ways
without persisting it, using RAFT~\cite{raft} running over TCP/IP. Either
persisting every update to the log or replicating it using TCP/IP causes
performance to drop by an order of magnitude (note the log scale on the
x-axis). It is therefore no suprise that deployments of Rocksdb often turn off
the write ahead log~\cite{samza}. On the other side of the spectrum, key value
store research such as FloDB~\cite{flodb} turns off the write ahead log to be
able to showcase benefits of sophisticated extensions to LSM datastructures.

@@

Cyclone is a high speed strongly consistent replicated write ahead logging
service specialized for key value stores such as Rocksdb. Cyclone provides the
order of magnitude better performance required to close the performance gap in
Figure~\ref{fig:problem} both on the storage and network side. Cyclone requires
\emph{a small amount of} directly attached Non Volatile Memory (NVM) on the
host - such as in the form of NVDIMMs~\cite{farm} or directly attached
persistent memory~\cite{pmfs}.

Cyclone makes use of a limited amount of attached NVM using two ideas. First, it
leverages the observation that with persistence being a first class property of
a subset of directly addressable DRAM, replicating the same log entry from the
leader to follower replicas is no different from multicast using software packet
switching techniques. It directly leverages this observation to implement a
version of RAFT that replicates the NVM log using classic packet switching
techniques such as separation of control and data planes. Second, Cyclone uses a
two level log structure where it drains entries from the NVM log in IO friendly
batches to a log placed on a flash drive. This allows Cyclone to use only a
limited amount of NVM following the reasoning that current forms of NVM such as
NVDIMMs are more expensive than DRAM. Our experiments use an NVM footprint of a
mere 64 MB to achieve performance comparable to placing the entire log in
NVM. This allows Cyclone to be used in deployments where the key value store is
maintained entirely in main memory rather than as a multi-level LSM tree
structure, with the log serving as the source for recovery in the event of
failure.

We discuss the two level structure focusing on the packet switching components
in Section~\ref{sec:2level}. We then describe how we can horizontally scale this
basic idea across multiple cores by leveraging the commutativity of key value
store APIs across different keys in Section~\ref{sec:horizontal}. We then
demonstrate in the evaluation in Section~\ref{sec:evaluation} that Cyclone can
replicate millions of updates a second using only commodity 10 Gigabit ethernet
adapters thereby closing the performance gap between key value store performance
and write ahead logging performance using only a small amount of directly
attached NVM. A discussion of related work follows before we conclude.

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.3]{figures2/network_packet.pdf}
  \caption{Cyclone network packet layout}
  \label{fig:packet_layout}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.3]{figures2/batching.pdf}
  \caption{Batching}
  \label{fig:batching}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.3]{figures2/flashlog_page.pdf}
  \caption{Flashlog segment}
  \label{fig:flashlog_page}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.3]{figures2/race.pdf}
  \caption{Race with multiple physical logs}
  \label{fig:race}
\end{figure}

>>>>

\documentclass[pageno]{jpaper}

%replace XXX with the submission number you are given from the ASPLOS submission site.
\newcommand{\asplossubmissionnumber}{XXX}

\usepackage[normalem]{ulem}

@@

\title{\Large \bf Cyclone: Warp Speed Replication for Key Value Stores}

\date{}
\maketitle

\thispagestyle{empty}

@@

\begin{abstract}
Persistent key value stores are rapidly becoming an important component of many
distributed data serving solutions with innovations targeted at taking advantage
of growing flash speeds. Unfortunately their performance is hampered by the need
to maintain and replicate a write ahead log to guarantee availability in the
face of machine and storage failures. Cyclone is a replicated log plug-in for
key value stores that systematically addresses various sources of this
bottleneck. It uses a small amount of non-volatile memory directly addressable
by the CPU - such as in the form of NVDIMMs - to remove block oriented IO
devices such as SSDs from the critical path for appending to the log. It then
addresses network overheads using an implementation of the RAFT consensus
protocol that is designed around a userspace network stack to relieve the CPU of
the burden of data copies. Finally, it provides a way to efficiently map the
commutativity in key-value store APIs to the parallelism available in commodity
NICs via multiple transmit receive queues to scale replication performance on
multicore machines. Cyclone is able to replicate millions of small updates a
second using only commodity 10 gigabit ethernet adapters. As a practical
application, we use it to improve the performance and availability of RocksDB, a
popular persistent key value store by an order of magnitude when compared to its
own write ahead log without replication.
\end{abstract}

@@

These key value stores however tend to ignore an important component: the write
ahead log. A machine or storage failure leading to total loss or temporary
unavailability of data is unacceptable in many internet scale services where
high availabilty and revenue are often interconnected. Key value stores
therefore usually incorporate support for a write ahead log that if replicated
and kept durable for every appended update provides the necessary high
availability. Unfortunately, the write ahead log is a performance achilles heel
for these systems - eclipsing much of the work on improving the performance of
the LSM component. To illustrate the impact of the write ahead log, consider
Figure~\ref{fig:problem}. The line marked 'Rocksdb' shows the performance of
Rocksdb without the write ahead log. The performance when persisting the write
ahead log without replicating it is shown as the line marked 'rocksdb/WAL'. The
line marked 'rocksdb/3 way rep.' is for simply replicating the log three ways
without persisting it, using RAFT~\cite{raft} running over TCP/IP. Either
persisting every update to the log or replicating it using TCP/IP causes
performance to drop by an order of magnitude (note the log scale on the
x-axis). It is therefore no surprise that deployments of Rocksdb often turn off
the write ahead log~\cite{samza}. On the other side of the spectrum, key value
store research prototypes such as FloDB~\cite{flodb} turns off the write ahead
log to be able to showcase benefits of sophisticated extensions to LSM
datastructures.

@@

Cyclone is a high speed strongly consistent replicated write ahead logging
service specialized for key value stores such as Rocksdb. Cyclone provides the
order of magnitude better performance required to close the performance gap in
Figure~\ref{fig:problem} both on the storage and network side. Cyclone requires
\emph{a small amount of} directly attached Non Volatile Memory (NVM) on the host
- such as in the form of NVDIMMs~\cite{farm} or directly attached persistent
memory~\cite{pmfs}. Cyclone logs update operations to the key value store and
calls on the key value store to execute the update only when it has been
successfully replicated to majority quorum of replicas. All reads and write to
the key value store are always executed by the leader replica. A client side
stub provided by Cyclone automatically locates and directs read and write
requests to the leader replica, ensuring linearizable semantics from the key
value store. Using a replicated write ahead log to build a linearizable and
highly available key value store is not new [find some examples here]. The
novelty in Cyclone is that it accomplishes this without compromising on the
performance baseline of no logging or replication. Cyclone does this by
systematically addressing the sources of performance degradation in
Figure~\ref{fig:problem}.

The primary cause of slowdown when turning on a write ahead log on a single
machine is the need to do synchronous block IO for every update. For even
moderately sized updates, this leads to severe write amplification as a 4KB page
must be synchronously written to the log file on flash. Cyclone avoids this by
making use of directly attached persistent memory on the server. This can be in
the form of NVDIMMs (DIMMs with an added ultracapacitor to dump state to a small
amount of attached flash) or newer forms of persistent memory~\cite{pmfs}. For
current deployments, NVDIMMs are expensive - retailing at approximately 2X the
price of a standard DIMM [find ref]. This means that they will likely be
available in small quantities. We compensate for this by periodically draining
the NVDIMM log into a log placed on a standard SSD on the system. We describe
this two level log structure in Section~\ref{sec:storage}.

The network overheads for replicating a log entry to a quorum of machines stems
from the overhead of the TCP/IP stack in the linux kernel - itself the topic of
much research [dataplane OS stuff]. We therefore replase the TCP/IP stack with
DPDK, a userspace network library that allows direct IO of raw ethernet frames
to and from the network interface card (NIC) eliminating all overheads. We no
longer use TCP - since the rare network errors encountered in a typical
datacenter network are automatically handled by the replication protocol (RAFT)
as part of its failure model. Finally, we carefully design our implementation of
RAFT to relieve the CPU of all data copy operations, leaving the core to only
run the protocol and return an accept/reject decision for each received
packet. The set of optimizations to the network stack is decribed in
Section~\ref{sec:network}.

Both the storage (NVM, SSD) as well as the NIC expose significant parallelism
that is left unused when Cyclone is used to replicate a single log. To exploit
this parallelism we use Cyclone to replicate multiple logs, each using a
different instance of the RAFT consensus protocol. We show how to exploit the
commutativity inherent in key-value store interfaces to make use of this
parallelism in Cyclone's design in Section~\ref{parallelism}.

A detailed evaluation of Cyclone's replication is provided in the context of
RocksDB - a popular persistent key value store in
Section~\ref{sec:evaluation}. We show that Cyclone achieves performance
transparent replication in that it can provide both a persistent write ahead log
and replication without compromising on RocksDB's performance.

\section{System Architecture}

\section{Storage}

\section{Network}

\section{Parallelism}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.35]{figures2/network_packet.pdf}
  \caption{Cyclone network packet layout}
  \label{fig:packet_layout}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.4]{figures2/batching.pdf}
  \caption{Batching}
  \label{fig:batching}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.4]{figures2/flashlog_page.pdf}
  \caption{Flashlog segment}
  \label{fig:flashlog_page}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.35]{figures2/race.pdf}
  \caption{Race with multiple physical logs}
  \label{fig:race}
\end{figure}
